---
title: "post1"
author: "Rongyun Tang"
date: "December 11, 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Census income 

research outlines:

1. Objective:task is to determine what kind of a person makes over 50K a year based on census data.

2. Method: statistics and association rules 

3. Data Source: this "Adult" dataset was downloaded from  UCI Machine Learning website: http://archive.ics.uci.edu/ml/datasets/Adult.  It is multivariated dataset(including categorical and Integer variables) from social area. A set of reasonably clean records was extracted by the data dornors.It is also splited into train-test using MLC++ GenCVFiles (2/3, 1/3 random).

=======================================================================================================

Attribute Information: 

=======================================================================================================
Listing of attributes: 
salaries are potentially divided into two classes: >50K, <=50K. 

age: continuous.
workclass: Private, Self-emp-not-inc, Self-emp-inc, Federal-gov, Local-gov, State-gov, Without-pay, Never-worked.
fnlwgt: continuous.
education: Bachelors, Some-college, 11th, HS-grad, Prof-school, Assoc-acdm, Assoc-voc, 9th, 7th-8th, 12th, Masters, 1st-4th, 10th, Doctorate, 5th-6th, Preschool.
education-num: continuous.
marital-status: Married-civ-spouse, Divorced, Never-married, Separated, Widowed, Married-spouse-absent, Married-AF-spouse.
occupation: Tech-support, Craft-repair, Other-service, Sales, Exec-managerial, Prof-specialty, Handlers-cleaners, Machine-op-inspct, Adm-clerical, Farming-fishing, Transport-moving, Priv-house-serv, Protective-serv, Armed-Forces.
relationship: Wife, Own-child, Husband, Not-in-family, Other-relative, Unmarried.
race: White, Asian-Pac-Islander, Amer-Indian-Eskimo, Other, Black.
sex: Female, Male.
capital-gain: continuous.
capital-loss: continuous.
hours-per-week: continuous.
native-country: United-States, Cambodia, England, Puerto-Rico, Canada, Germany, Outlying-US(Guam-USVI-etc), India, Japan, Greece, South, China, Cuba, Iran, Honduras, Philippines, Italy, Poland, Jamaica, Vietnam, Mexico, Portugal, Ireland, France, Dominican-Republic, Laos, Ecuador, Taiwan, Haiti, Columbia, Hungary, Guatemala, Nicaragua, Scotland, Thailand, Yugoslavia, El-Salvador, Trinadad&Tobago, Peru, Hong, Holand-Netherlands.


```{r initialization}
## 1. Data preprocessing
training.data <- as.data.frame(read.csv('adult.data'))
test.data <- as.data.frame(read.csv('adult.test', skip=1))
content <- readLines('old.adult.names')

i <- grep("Attribute Information",content) + 2
  var.names <- NULL
  while(content[i]!="") {
    j <- gregexpr(":", content[i])[[1]][1]
    var.names <- c(var.names, substr(content[i],1,j-1))
    i <- i + 1
  }
  names(training.data) <- gsub("-","",var.names)
  names(test.data) <- gsub("-","",var.names)
  N.obs <- dim(training.data)[1]  
  N.var <- dim(training.data)[2]
  
  # show some data information: 
  print("Traning data examples: ")
  head(training.data) 
  cat("Number of observations:",N.obs,"

     Number of variables:",N.var,"\n")
```


```{r data preprocessing: missing values detection}
#install.packages('VIM')
library("VIM")

# missing values detection 
print("This is for missing values detection:")
aggr(training.data,prop=FALSE,numbers=TRUE)
aggr(test.data,prop=FALSE,numbers=TRUE)
print("Numbers of missing values in trainning data and test dat: 0 , 0 ")

# since most of the variables are categorial, so we didn't do outliers detection here 

```
## 2. statistics on numerical variables 

```{r numerical data distribution in two groups }
par(mfrow=c(2,2))  ## Arrange plots in a 4x4 grid
boxplot(training.data[,'age']~training.data[,'class'], main="Age vs. Income Class", 
        xlab="Income Class", ylab="Age")
boxplot(training.data[,'educationnum']~training.data[,'class'], main="Years of Eduction vs. Income Class", 
        xlab="Income Class", ylab="Years of Eduction")
boxplot(log(training.data[,'fnlwgt'])~training.data[,'class'], main="log(Weight) vs. Income Class", 
        xlab="Income Class", ylab="log(Weight)")
boxplot(training.data[,'hoursperweek']~training.data[,'class'], main="Hours per Week vs. Income Class", 
        xlab="Income Class", ylab="Hours per Week")
par(mfrow=c(1,1))
```


numerical data distribution analysis: 

1. In group that income >50k, people are more likely to have higher average age, higher average education years and higher weekly work hours than whose income is <=50K. 

2. Weight shows no difference in these two classes. 

## 3. association rules of categorial variables 

```{r  apply association rule on categorial variables of traning data }
library(arules)
training<-training.data[,c(2,4,6:10,14,15)]
summary(training)
# use apriori rules to find association rules on people whose income class is >50K 
rules <- apriori(training,
                 control = list(verbose=F),
                 parameter = list(minlen=2, supp=0.005, conf=0.8),
                 appearance = list(rhs=c("class= >50K"),
                                   default="lhs"))

inspect(sort(rules, by="lift", decreasing = TRUE)[1:5])
```
## 4. association rules validation

```{r  association rule validatiion on testing data}
# using test data and comfusion matrix to test the results 
test<-test.data[,c(2,4,6:10,14,15)]
summary(test)

test.rule1<-subset(test, (workclass==' Private') & (education==' Masters') & (occupation==' Exec-managerial') & (relationship==' Husband') & (nativecountry==' United-States'))
accuracy.rules1=nrow(subset(test.rule1,class==' >50K.'))/nrow(test.rule1)

test.rule2<-subset(test,workclass==' Private' & education==' Masters' & maritalstatus==' Married-civ-spouse' & occupation==' Exec-managerial' & relationship==' Husband' & nativecountry==' United-States')
accuracy.rules2=nrow(subset(test.rule2,class==' >50K.'))/nrow(test.rule2)

test.rule3<-subset(test,workclass==' Private' & education==' Masters' & occupation==' Exec-managerial' & relationship==' Husband' & sex==' Male')
accuracy.rules3=nrow(subset(test.rule3,class==' >50K.'))/nrow(test.rule3)


test.rule4<-subset(test,workclass==' Private' & education==' Masters' & maritalstatus==' Married-civ-spouse' & occupation==' Exec-managerial' & sex==' Male' & nativecountry==' United-States')
accuracy.rules4=nrow(subset(test.rule4,class==' >50K.'))/nrow(test.rule4)


test.rule5<-subset(test,workclass==' Private' & education==' Masters' & maritalstatus==' Married-civ-spouse' & occupation==' Exec-managerial' & relationship==' Husband' & sex==' Male' & nativecountry==' United-States')
accuracy.rules5=nrow(subset(test.rule5,class==' >50K.'))/nrow(test.rule5)

paste('accuracy of association rule1,rule2,rule3,rule4 and rule5 are:',round(accuracy.rules1,3),round(accuracy.rules2,3),round(accuracy.rules3,3), round(accuracy.rules4,3), round(accuracy.rules5,3))

```

## 5. conclusions

```{r}
#1. numerical statistic analysis showed that people have higher average age, higher average education years and higher weekly work hours are more likely to have income >50K. 

#2. association rules listed in this project showed good accuracies: rule1(91%),rule2(91%),rule3(89%),rule4(91%),rule5(91%). 

#3. association rules showed that people who have master degrees, work privately married male with an Exec-managerial occupation have large possibilities(>80%) to have salaries more than 50,000. 
```



