---
title: "post3-PCA"
author: "Rongyun Tang"
date: "December 11, 2018"
output: 
       rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Daily Demand Forecasting Orders Data  

research outlines:

1. Objective:task is to determine what kind of order resources contribute most to the information, and then to reduce data dimentions. 

2. Method: CPA 

3. Data Source:this "Daily Demand Forecasting Orders Data(DDFOD)" dataset was downloaded from  UCI Machine Learning website: https://archive.ics.uci.edu/ml/datasets/Daily+Demand+Forecasting+Orders.
 The dataset was collected during 60 days, this is a real database of a brazilian logistics company.Twelve predictive attributes and a target that is the total of orders for daily treatment.

===================================================================================

Attribute Information: 

===================================================================================

The dataset was collected during 60 days, this is a real database of a brazilian logistics company. The dataset has twelve predictive attributes and a target that is the total of orders for daily treatment. 

- Week_of_the_month (WM): {1.0, 2.0, 3.0, 4.0, 5.0}
- Day_of_the_week_(Monday_to_Friday)(DW): {2.0, 3.0, 4.0, 5.0, 6.0}
- Non_urgent_order(NUO): integer
- Urgent order(UO): integer
- Order type A(typeA): integer
- Order type B(typeB): integer
- Order type C(typeC):integer
- Fiscal sector orders(FO): integer
- Orders from the traffic controller sector(Traffic): integer
- Banking orders (1)(Bank1): integer
- Banking orders (2)(Bank2): integer
- Banking orders (3)(Bank3): integer
- Target(Total_orders)(Total):integer

## 1. Data Preprocessing 

```{r data preprocessing}
# Rename column names and Read data
rawdata <- read.csv('Daily_Demand_Forecasting_Orders2.csv')
head(rawdata)
DOY<-as.factor(rawdata$WM*30+rawdata$DW)    # transfer week of month and day of week into day of year(DOY)
data2<-cbind(DOY,rawdata[,3:13])
head(data2)

```

## 2.Data Exploration

```{r numerial data exploration}

# Missing value detection 
sum(is.na(data2))
print("Numbers of missing values : 0")


# Data Correlationships

#install.packages("ggcorrplot")
library('ggcorrplot')
corr <- round(cor(data2[,c(-1,-12)]), 1)
head(corr[, 1:6])
ggcorrplot(corr, hc.order = TRUE, type = "lower",lab = TRUE)

# Resutls show that there are many variables mutually correlated, dimension could be reduced with PCA method.
```

## 3. PCA and Visualization

```{r }
# 3.1 apply PCA

library(factoextra)
data.pca <- prcomp(data2[,c(-1,-12)], center = TRUE,scale. = TRUE)
summary(data.pca)

# 3.12 results: PC1 explains 44% of the total variance,  PC2 explains 16% of the variance , PC3 expalins 15% of the total variance

fviz_eig(data.pca)

#3.2 Individuals with a similar profile are grouped together
fviz_pca_ind(data.pca,  
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )

library(devtools)
library(ggbiplot)
ggbiplot(data.pca, labels=colnames(data2))


# 3.3 Variables with a similar profile are grouped together
#     Positive correlated variables point to the same side of the plot. Negative correlated variables point to o           opposite sides of the graph.
fviz_pca_var(data.pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
# 3.32 results: NUO(not urgent orders) and Bank1 are two most strong positive-correlated variables. Bank3 and FO have opposite correlationship. NUO,Bank1 and typeC are three main contributorS to PAC. 

# 3.4 Biplot of individuals and variables
fviz_pca_biplot(data.pca, repel = TRUE,
                col.var = "#2E9FDF", # Variables color
                col.ind = "#696969"  # Individuals color
                )
```

## 4. PCA results

```{r PCA results}
library(factoextra)
# Eigenvalues
eig.val <- get_eigenvalue(data.pca)
eig.val
  
# Results for Variables
res.var <- get_pca_var(data.pca)
head(res.var$coord)          # Coordinates
head(res.var$contrib)        # Contributions to the PCs
head(res.var$cos2)  # Quality of representation 
library("corrplot")
corrplot(res.var$cos2, is.corr=FALSE)

# Results for individuals
res.ind <- get_pca_ind(data.pca)
head(res.ind$coord)         # Coordinates
head(res.ind$contrib)        # Contributions to the PCs
head(res.ind$cos2)          # Quality of representation 
corrplot(res.ind$cos2,is.corr=FALSE)

```



## 5. Conclusion


- 13 variables of the raw data could be reduced by 2 PCA principles. 

- NUO contributes most to the effective information,  followed by type B and Bank2.

- individual comtribution shows that types C has seasonal attribute, and it reaches maxmiu contribution in April. 



